|

### 2023-04-09 paper.Rmd
abb BP and SW

We can assess residuals in multiple ways. Residuals may be plotted as a histogram or quantile-quantile plot to examine the distribution. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it to be well behaved. However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern. One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk Normality test [@shapiro1965analysis].

More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships. If there are any visually discoverable patterns, the model is potentially misspecified. In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans. It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing]. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan test [@breusch_simple_1979] for heteroskedasticity.

@@

Many different hypothesis tests are available to detect specific model defects. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the Breusch-Pagan test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969]. The Shapiro-Wilk test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs. The Jarque--Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.

Table \ref{tab:example-residual-plot-table} displays the $p$-values from the RESET, Breusch-Pagan and Shapiro-Wilk tests applied to the residual plots in Figure \ref{fig:residual-plot-common-departures}. The RESET test and Breusch-Pagan test were computed using the `resettest` and `bptest` functions from the R package `lmtest`, respectively. The Shapiro-Wilk test was computed using the `shapiro.test` from the core R package `stats`. (The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) Although, the RESET test is exact, it requires the selection of a power parameter. According to @ramsey_tests_1969, a power of four is recommended, which we adopted in our analysis.  The Breusch-Pagan and Shapiro-Wilk tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot B, the Breusch-Pagan test for heteroskedasticity to reject the residual plot C, and Shapiro-Wilk test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A. Interestingly, the Breusch-Pagan and Shapiro-Wilk tests also reject the residual plots exhibiting structure that they weren't designed for. @cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error". Also, some types of departure can also have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity. Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].

@@

<!--
As discussed in Section \ref{conventionally-testing-for-departures}, many conventional tests are available for detecting residual departures. Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as Shapiro-Wilk test. A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest]. In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.

We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the Breusch-Pagan test be the same as the White test. Thus, we include both $x$ and $x^2$ in the auxiliary regression.
-->

@@

  table_dat %>%
  knitr::kable(col.names = c("Plot", 
                             "Departures", 
                             "RESET", 
                             "Breusch-Pagan", 
                             "Shapiro–Wilk"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the RESET, the Breusch-Pagan and the Shapiro–Wilk tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" ) 
```

@@

However, it is practically impossible to test this specific $H_0$ with conventional tests, which are constructed to measure specific departures. For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$. 
Similarly, the Breusch-Pagan test is designed to specifically test $H_0:$ *error variances are all equal* ($\zeta_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \zeta_i\neq 0$) from $e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)$.

@@

Figures \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals. The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity. The power curves for the visual test are effectively a shift right from that of the RESET test. This means that the RESET test will reject a lower effect size (less structure) than the visual test, but otherwise the performance will be similar. In other words, the RESET test is more sensitive than the visual test. This is not necessarily a good feature, for the purposes of diagnosing model defects. If one scans the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5. The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25. As expected the Breusch-Pagan and Shapiro-Wilk tests have much lower power - they are not designed to detect non-linearity.

For the heteroskedasticity pattern, the power of Breusch-Pagan test, designed for detecting heteroskedasticity, is uniformly higher than the other tests. The visual test power curve is a right shift. This shows a similar story to the power curves for non-linearity pattern - the conventional test is more sensitive than the visual test. From the example residual plots at the bottom we might argue that the heteroskedasticity becomes noticeably visible around an effect size of 3 or 3.5. However the Breusch-Pagan test would reject at around effect size 2.5. Interestingly, the power curve for the Shapiro-Wilk test (for non-normality) is only sighly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too. The power curve for the Breusch-Pagan test suggests it is not useful for detecting heteroskeadsticity, as expected.

@@

Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject. Figure \ref{fig:heter-example} shows this lineup. The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$). This is reflected by the visual test $p\text{-value} = 0.026$. But the Breusch-Pagan test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$. This lineup was evaluated by 11 subjects, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values. It may have been the small sample size and the presence of a few outliers that may have resulted in the lack of detection by the conventional test.

@@

```{r different-x-dist-poly-power, fig.pos="t!", fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns. The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and Breusch-Pagan tests, respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (grey). Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests. Uneven distributions, normal and lognormal distributions, tend to yield lower power.", dev = 'png', dpi = 450}

@@

Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional (top row) and visual (bottom row) tests for both the non-linearity (left column) and heteroskedasticity (right column) patterns. For conventional tests, only the power curves of appropriate tests which are RESET tests and Breusch-Pagan tests for non-linearity pattern and heteroskedasticity pattern respectively are shown. For visual tests, note that, more evaluations on lineup with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the total eleven evaluations to estimate the power curves. Re-sampling produces the multiple curves for the uniform, and provides an indication of the variation of the power.

>>>>

We can assess residuals in multiple ways. Residuals may be plotted as a histogram or quantile-quantile plot to examine the distribution. Using the classical normal linear regression model as an example, if the distribution is symmetric and unimodal, we consider it to be well behaved. However, if the distribution is skewed, bimodal, multimodal, or contains outliers, there is a cause for concern. One could also inspect the distribution by conducting a goodness of fit test, such as the Shapiro-Wilk (SW) Normality test [@shapiro1965analysis].

More typically, residuals will be plotted, as a scatter plot against the predicted values and each of the explanatory variables to scrutinize their relationships. If there are any visually discoverable patterns, the model is potentially misspecified. In general, one looks for noticeable departures from the model like non-linear dependency or heteroskedasticity. However, correctly judging a residual plot where no pattern exists can be a painstakingly difficult task for humans. It is especially common, particularly among students, to misinterpret patterns that are random noise and random deviation from a model [@loy2021bringing]. It is also possible to conduct hypothesis tests for non-linear dependence [@ramsey_tests_1969], and use a Breusch-Pagan (BP) test [@breusch_simple_1979] for heteroskedasticity.

@@

Many different hypothesis tests are available to detect specific model defects. For example, the presence of heteroskedasticity can usually be tested by applying the White test [@white_heteroskedasticity-consistent_1980] or the BP test [@breusch_simple_1979], which are both derived from the Lagrange multiplier test [@silvey1959lagrangian] principle that relies on the asymptotic properties of the null distribution. To test specific forms of non-linearity, one may apply the F-test as a model structural test to examine the significance of specific polynomial and non-linear forms of the predictors, or the significance of proxy variables as in the Ramsey Regression Equation Specification Error Test (RESET) [@ramsey_tests_1969]. The SW test [@shapiro1965analysis] is the most widely used test of non-normality included by many of the statistical software programs. The Jarque--Bera test [@jarque1980efficient] is also used to directly check whether the sample skewness and kurtosis match a normal distribution.

Table \ref{tab:example-residual-plot-table} displays the $p$-values from the RESET, BP and SW tests applied to the residual plots in Figure \ref{fig:residual-plot-common-departures}. The RESET test and BP test were computed using the `resettest` and `bptest` functions from the R package `lmtest`, respectively. The SW test was computed using the `shapiro.test` from the core R package `stats`. (The the R package `skedastic` [@skedastic] contains a large collection of tests for heteroskedasticity.) Although, the RESET test is exact, it requires the selection of a power parameter. According to @ramsey_tests_1969, a power of four is recommended, which we adopted in our analysis.  The BP and SW tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot B, the BP test for heteroskedasticity to reject the residual plot C, and SW test for non-normality to reject residual plot D, which they all do and correctly fail to reject residual plot A. Interestingly, the BP and SW tests also reject the residual plots exhibiting structure that they weren't designed for. @cook1982residuals discusses that most residual-based tests for particular types of departure from model assumptions are also sensitive to other types of departures, that is, correctly rejected but for the wrong reason, a phenomenon known as the "Type III error". Also, some types of departure can also have elements of other types of departure, for example, non-linearity could be viewed as heteroskedasticity. Additionally, other problesms such as outliers can trigger the rejection [@cook_applied_1999].

@@

<!--
As discussed in Section \ref{conventionally-testing-for-departures}, many conventional tests are available for detecting residual departures. Implementation-wise, the built-in R package `stats` provides some commonly used residual-based tests, such as SW test. A more comprehensive collection of regression diagnostics tests can be found in the R package `lmtest` [@lmtest]. In terms of heteroskedasticity diagnostics, the R package `skedastic` [@skedastic] collects and implements 25 existing conventional tests published since 1961.

We pick RESET test (`resettest`) and BP test (`bptest`) from the R package `lmtest`, and SW test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the BP test and the SW test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the BP test, the choice of predictors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary predictors in the BP test be the same as the White test. Thus, we include both $x$ and $x^2$ in the auxiliary regression.
-->

@@

  table_dat %>%
  knitr::kable(col.names = c("Plot", 
                             "Departures", 
                             "RESET", 
                             "BP", 
                             "SW"),
               align = "llrrr",
               escape = FALSE,
               format = "latex",
               booktabs = TRUE, 
               caption  = 'Statistical significance testing for departures from good residuals for plots in Figure \\ref{fig:residual-plot-common-departures}. Shown are the $p$-values calculated for the RESET, the BP and the SW tests. The good residual plot (A) is judged a good residual plot, as expected, by all tests. The non-linearity (B) is detected by all tests, as might be expected given the extreme structure.', 
               label = "example-residual-plot-table" ) 
```

@@

However, it is practically impossible to test this specific $H_0$ with conventional tests, which are constructed to measure specific departures. For example, the RESET test is formulated as $H_0:\gamma_1 = \gamma_2 = \gamma_3 = 0$ against $H_a: \gamma_1 \neq 0 \text{ or } \gamma_2 \neq 0 \text{ or } \gamma_3 \neq 0$, from $y = \tau_0 + \sum_{i=1}^{p}\tau_px_p +\gamma_1\hat{y}^2 + \gamma_2\hat{y}^3 + \gamma_3\hat{y}^4 + u, ~~u \sim N(0, \sigma_u^2)$. 
Similarly, the BP test is designed to specifically test $H_0:$ *error variances are all equal* ($\zeta_i=0 \text{ for } i=1,..,p$) versus the alternative $H_a:$ *that the error variances are a multiplicative function of one or more variables* ($\text{at least one } \zeta_i\neq 0$) from $e^2 = \zeta_0 + \sum_{i=1}^{p}\zeta_i x_i + u, ~ u\sim N(0,\sigma_u^2)$.

@@

Figures \ref{fig:polypower} compares the power for the different tests for non-linear structure in the residuals. The test with the uniformly higher power is the RESET test, one that specifically tests for non-linearity. The power curves for the visual test are effectively a shift right from that of the RESET test. This means that the RESET test will reject a lower effect size (less structure) than the visual test, but otherwise the performance will be similar. In other words, the RESET test is more sensitive than the visual test. This is not necessarily a good feature, for the purposes of diagnosing model defects. If one scans the residual plot examples at the bottom, we might argue that the non-linearity is not sufficiently problematic until an effect size of around 3 or 3.5. The RESET test would reject closer to an effect size of 2, but the visual test would reject closer 3.25. As expected the BP and SW tests have much lower power - they are not designed to detect non-linearity.

For the heteroskedasticity pattern, the power of BP test, designed for detecting heteroskedasticity, is uniformly higher than the other tests. The visual test power curve is a right shift. This shows a similar story to the power curves for non-linearity pattern - the conventional test is more sensitive than the visual test. From the example residual plots at the bottom we might argue that the heteroskedasticity becomes noticeably visible around an effect size of 3 or 3.5. However the BP test would reject at around effect size 2.5. Interestingly, the power curve for the SW test (for non-normality) is only sighly different to that of the visual test, suggesting that it performs reasonably for detecting heteroskedasticity, too. The power curve for the BP test suggests it is not useful for detecting heteroskeadsticity, as expected.

@@

Surprisingly, the visual test rejects `r nr$n[1]` of the `r nr$n[1]+nr$n[2]` (`r round(nr$n[1]/(nr$n[1]+nr$n[2])*100, 0)`%) of lineups where the conventional test does not reject. Figure \ref{fig:heter-example} shows this lineup. The data plot in position seventeen displays a relatively strong heteroskedasticity pattern, and has a strong effect size ($log_e(E)=4.02$). This is reflected by the visual test $p\text{-value} = 0.026$. But the BP test $p\text{-value} = 0.056$, is slightly above the significance cutoff of $0.05$. This lineup was evaluated by 11 subjects, it has experimental factors $a = 0$ ("butterfly" shape), $b = 64$ (large variance ratio), $n = 50$ (small sample size), and a uniform distribution for the fitted values. It may have been the small sample size and the presence of a few outliers that may have resulted in the lack of detection by the conventional test.

@@

```{r different-x-dist-poly-power, fig.pos="t!", fig.height = 6, fig.cap = "Comparison of power on lineups with different fitted value distributions for conventional and visual tests for non-linearity and heteroskedasticity patterns. The power curves of conventional tests for non-linearity and heteroskedasticity patterns are produced by RESET tests and BP tests, respectively. Power curves of visual tests are estimated using five evaluations on each lineup. For lineups with a uniform fitted value distribution, the five evaluations are repeatedly sampled from the total eleven evaluations to give multiple power curves (grey). Surprisingly, the fitted value distribution has produces more variability in the power of conventional tests than visual tests. Uneven distributions, normal and lognormal distributions, tend to yield lower power.", dev = 'png', dpi = 450}

@@

Figure \ref{fig:different-x-dist-poly-power} examines the impact of the fitted value distribution on the power of conventional (top row) and visual (bottom row) tests for both the non-linearity (left column) and heteroskedasticity (right column) patterns. For conventional tests, only the power curves of appropriate tests which are RESET tests and BP tests for non-linearity pattern and heteroskedasticity pattern respectively are shown. For visual tests, note that, more evaluations on lineup with uniform fitted value distribution were collected, so to have a fair comparison, we randomly sample five from the total eleven evaluations to estimate the power curves. Re-sampling produces the multiple curves for the uniform, and provides an indication of the variation of the power.
